---
title: "Data 605 - Discussion Board12"
author: "IvanTikhonov"
date: "2022-11-09"
output: html_document
---
Multiple Regression

Real estate price prediction:Regression analysis, linear regression, multiple regression,
and prediction models were all used to build this real estate dataset.
It covers the purchase date, the age of the house, the location,
the distance to the closest MRT station, and the housing price per square foot.

Data set was acquired using https://www.kaggle.com/datasets/quantbruce/real-estate-price-prediction

```{r}
# Load Libraries
library(tidyverse)
```

```{r}
# Import data
real_estate <- read.csv("https://raw.githubusercontent.com/IvanGrozny88/DATA605Assignment_12/main/Real%20estate.csv")
head(real_estate, n = 5)
```
This dataset has no missing values.

```{r}
# Checking is missing values
head(is.na(real_estate))
```
There are 414 rows and 8 columns in this data set.

```{r}
# Summary of data set
summary(real_estate)
```
```{r}
# Getting colnames 
colnames(real_estate)
```
```{r}
# Renaming for easier read
colnames(real_estate) <- c("No", "Transaction_Date", "House_Age", 
                           "Nearest_MRT_Station", "Number_Convenience_Stores",
                           "Latitude", "Longitude", "House_Price")
head(real_estate)
```

Regression Analysis
I started by running a simple linear model using the variables house price and house age.
The multiple R2 is 0.4434, which according to the summary explains 44.34% of the data model's variance,
and has a p-value of 1.56e-05, which is less than 0.05.

```{r}
# Obtaining our singular Linear Model
my_lm <- lm(House_Price ~ House_Age, real_estate)
summary(my_lm)
```
Managing various regressions To see how the two models would differ from one another,
I developed a second model and this time added extra coefficients.
There are already some variations between the multiple, coefficients, and residuals.
The current R2 is 58.24%, and the p-value is 2.2e-16.

```{r}
# Running the multiple regression
my_lm2 <- lm(House_Price ~ Transaction_Date + House_Age +  
                           Nearest_MRT_Station + Number_Convenience_Stores +
                           Latitude + Longitude, real_estate)
summary(my_lm2)
```
Since Longitude has a p-value above 0.05, as can be seen,
I will eliminate it to create a third regression model.
The coefficient values below show some tiny variations,
but they are all less than 0.05.

```{r}
# Multiple regression without Longitude
my_lm3 <- lm(House_Price ~ Transaction_Date + House_Age +  Nearest_MRT_Station + 
               Number_Convenience_Stores + Latitude, real_estate)
summary(my_lm3)
```
Residual Analysis

```{r}
# my_lm
par(mfrow = c(2,2))
plot(my_lm)
```
```{r}
# my_lm2
par(mfrow = c(2,2))
plot(my_lm2)
```

```{r}
# my_lm3
par(mfrow = c(2,2))
plot(my_lm3)
```
```{r}
# my_lm
hist(my_lm$residuals, xlab = 'Residuals', main = 'Histogram of Singluar Linear Model')
```
```{r}
#my_lm2
hist(my_lm2$residuals, xlab = 'Residuals', main = 'Histogram of Multiple Linear Model')
```
```{r}
#my_lm3
hist(my_lm3$residuals, xlab = 'Residuals', main = 'Histogram of Multiple Linear Model without Longitude')
```
Conclusion:
This model does fit the data because residuals are regularly distributed,
particularly for the second and third models that were developed.



